There are many people who in one form or another contributed to the work
described in this thesis. 

I would like to thank Ingo Sander, my supervisor, for his flexibility
and trust during the thesis.  He has always been positive about my,
sometimes not-so-frequent and too extensive, thesis status reports. I
would also like thank him and give him credit for kindly letting me
use some of the figures of his PhD thesis \cite{forsyde:thesis}.

I am also tremendously thankful to the Haskell community, which has proven to
be, by far, the most friendly, helpful and smart programming community I have
ever been involved with. Special thanks go to the guys at the
\texttt{\#haskell} IRC channel at \texttt{freenode.net} and the
\texttt{haskell-cafe} mailing list. There are too many community-people to
thank but I would like to specifically mention Oleg Kiselyov for his selfless
help fixing and designing the compiler's typechecker.

Koen Claessen gave me access to Lava's sourcecode, which turned to be essential
during the developmnent of the compiler. Without Lava, the compiler would
have been totally different and perhaps even unfeasible.

Last but not least, I would like to thank Iv\'{a}n P\'{e}rez and
Miguel Jim\'{e}nez who helped reviewing parts of the thesis, my
parents, for their support, whithout which I would not had been able
to spend my last two academic years in Stockholm and finally, all the
Erasmus students which turned my stay in Sweden into one of the best
experiences of my life (they know who they are).

\vspace{\baselineskip} 
\noindent Stockholm, \thedate

\vspace{\baselineskip} 
\noindent\theauthor